---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CaribouDemographyBasicApp

<!-- badges: start -->
<!-- badges: end -->

```{r, echo=FALSE, results='asis'}
inst_dir <- system.file(package = "CaribouDemographyBasicApp")
htmltools::includeMarkdown(file.path(inst_dir, "app_text", "intro_en.md")) |>
  stringr::str_replace("www/Female_caribou_and_calves.jpg",
                       file.path(inst_dir, "www/Female_caribou_and_calves.jpg"))|>
  knitr::raw_html()
```


## Installation

There are two options available for installing the app. Users who do not have R installed and are not familiar with R packages can follow the instructions on the [Releases](https://github.com/LandSciTech/CaribouDemographyBasicApp/releases) page to download a version of the app that is packaged with a version of R. Please note, you will not be able to update the caribou demographic data stored with the app when it is installed this way.

R users or those who wish to update the caribou demographic data used in the app can follow the instructions below to install the R package.  

To run the app you will need to install R and we recommend RStudio as well.
Follow the instructions [here](https://posit.co/download/rstudio-desktop/) to
install both.

In addition, you will need to install the Rtools application. You can do this by selecting the version corresponding to your version of R [here](https://cran.r-project.org/bin/windows/Rtools/). Or you can run the following code in the R console which should prompt Rtools to be installed. 

```r
install.packages("pkgbuild")
pkgbuild::check_build_tools()

```
This will check if you already have Rtools installed and if not it will ask you if you would like to install it. Wait for the tool to download and install, following the instructions that pop-up. When installation is complete close R or RStudio and open it again in a fresh session for the changes to take effect. You can run `pkgbuild::check_build_tools()` again to confirm that the installation was successful. 

You can install the CaribouDemographyBasicApp R package by running the following
in the R console:

``` r
install.packages("remotes")
remotes::install_github("LandSciTech/CaribouDemographyBasicApp")

```
Note this may take a long time the first time you install it since it will also install all the R packages that CaribouDemographyBasicApp depends on. 

## Launching the Shiny app

To launch the app in your default browser run the following lines:

``` r
library(CaribouDemographyBasicApp)
run_caribou_demog_app()

```

To launch the app in French include the lang = "fr" argument:

``` r
run_caribou_demog_app(lang = "fr")

```


## Updating the R package and app

If you want to see the latest changes made to the app you will need to re-install 
the package from GitHub. Run the code below in the R console:

``` r
remotes::install_github("LandSciTech/CaribouDemographyBasicApp",
                        type = "binary", upgrade = "always")

```
This will install the most recent version of the app, upgrade any dependencies that have updates available and it will only use binary package versions. This avoids the need for any user input during the installation. 

