---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CaribouDemographyBasicApp

<!-- badges: start -->
<!-- badges: end -->

```{r, echo=FALSE}
inst_dir <- system.file(package = "CaribouDemographyBasicApp")
htmltools::includeMarkdown(file.path(inst_dir, "app_text", "intro_en.md"))
```


## Installation

To run the app you will need to install R and we recommend RStudio as well.
Follow the instructions [here](https://posit.co/download/rstudio-desktop/) to
install both.

In addition you will need to install the Rtools application. You can do this by selecting the version corresponding to your version of R [here](https://cran.r-project.org/bin/windows/Rtools/). Or you can run the following code in the R console which should prompt Rtools to be installed. 

```r
install.packages("pkgbuild")
pkgbuild::check_build_tools()

```
Wait for the tool to download and install, following in instructions that pop-up. When installation is complete close R or RStudio and open it again in a fresh session for the changes to take effect. 

You can install the CaribouDemographyBasicApp R package by running the following
in the R console:

``` r
install.packages("remotes")
remotes::install_github("LandSciTech/CaribouDemographyBasicApp")

```

## Launching the Shiny app

To launch the app in your default browser run the following lines:

``` r
library(CaribouDemographyBasicApp)
run_caribou_demog_app()

```

To launch the app in French include the lang = "fr" argument:

``` r
run_caribou_demog_app(lang = "fr")

```


## Updating the R package and app

If you want to see the latest changes made to the app you will need to re-install 
the package from GitHub. Run the code below in the R console:

``` r
remotes::install_github("LandSciTech/CaribouDemographyBasicApp", upgrade = FALSE)

```

